---
title: "Class-based network segregation, Economic Inequality and Redistributive Preferences across societies"
css: "input/css/custom.css" # custom CSS para html
fontsize: 12pt
linestretch: '1.15'          # interlineado 
papersize: a4
link-citations: yes         # citas linkeadas
lang: en-US
# date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
date: "`r format(Sys.Date(), '%d %b %Y')`"
author:
- name: Julio Iturra-Sanhueza
#   affiliation: University of Bremen
#   email: jiturra@uni-bremen.de
#   number: 1
# - name: Justin d'Ottawa
#   affiliation: University of Ottawa
#   number: 2
# - name: Pedro Torres
#   number: 1,2 
# Nota: Autores que comparten filiacion, poner el mismo number y declarar filiación una sola vez.  
# abstract: |
#  The escalating economic inequality and the current sanitary crisis have threatened social cohesion among citizens. However, these consequences have been unevenly experienced across the social structure. The upper and intermediate classes, benefiting from privileged access to social resources, have shown resilience. Conversely, working-class families have faced deteriorating material conditions, leading to a heightened sense of marginalisation and increasing the demand for welfare support. Particularly, social resources through network ties have been suggested to have a direct link with an individual’s welfare through instrumental and expressive outcomes, such as providing information or help in moments of need. Consequently, inequalities in access to certain social positions (i.e. occupations) are translated into a lack of social resources beyond individual economic and cultural capital. Additionally, it has been argued that social isolation not only plays a role in terms of resources but given that higher levels of segregation can bolster opinions, it is argued that being isolated from other social classes can polarise attitudes mainly in the working and upper-middle classes. Employing data from the International Social Survey Programme - Social Networks  this research aims to understand to what extent the demand for redistribution is linked to social ties. To addresss this, ego-centred social networks are used to represent the degree of class-based homogeneity of social ties. The results of multilevel estimations show that being socially segregated increases demand for redistribution. Additionally, this influence is conditional to social class, where the working and intermediate classes with highly homogeneous networks demand more governmental redistribution than the upper class. At the macro level, as the current distribution of economic resources influences the opportunity structure, it has been hypothesised that income inequality could bolster the influence of social segregation on the demand for redistribution. However, cross-level interactions do not provide supporting evidence on this behalf. 
#    **Keywords**: network segregation, redistributive preferences, economic inequality 
output:
  bookdown::pdf_document2:
    template: null
    toc: false
    keep_tex: false
    number_sections: true
    pandoc_args:
      - --template=input/mytemplate.tex #custom template para usar autores con afiliacion  
  bookdown::html_document2:
    toc: yes
    code_folding: hide
    number_sections: false
    toc_float:
      collapsed: false
      smooth_scroll: false
         
always_allow_html: true      
linkcolor: black                         # enlaces y citas en color azul
bibliography: input/bib/rgroup-redist.bib     # bibliografia en bibtex
csl: input/bib/apa6.csl
editor_options:
  chunk_output_type: console            # en RStudio, mostrar output en consola
geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm" # márgenes de página
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \setlength\parindent{24pt} 
  - \usepackage{fancyhdr}
  - \fancyhead{} 
---

```{r eval=FALSE, include=FALSE}
 # for render in pdf run rmarkdown::render_site("docs/paper.Rmd", output_format = "all")
 # clean #in the yml
 rmarkdown::render("paper1_jiturra.Rmd", output_format = "bookdown::pdf_document2")
 rmarkdown::render("paper1_jiturra.Rmd", output_format = "bookdown::html_document2")
 rmarkdown::render("paper1_jiturra.Rmd", output_format = "bookdown::word_document2")
```


<!-- - considerar ISEI08 como alternativa de social class por la correlacion con homogeneidad -->
<!-- - darle una oportunidad al h-index -->
<!--     - en general es prácticamente lo mismo que el indicador de homogeneidad -->
<!-- - controlaria por social activities index () -->
<!-- - social resources  -->
    
```{r setup, include=FALSE}
if (!require("pacman")) install.packages("pacman")  #si falta pacman, instalar
if (!require("tinytex")) install.packages("tinytex")#si falta tinytex, instalar
pacman::p_load(knitr, kableExtra, dplyr, ggplot2,sjmisc,texreg) # librerias
knitr::opts_chunk$set(warning = FALSE,  # mensaje de warning
                      message = FALSE,  # mensajes/avisos de librerias  
                      cache = F,    # cache de los chunks,usar analisis pesados
                      out.width = '100%',
                      fig.pos= "H",     # posicion figuras H = HERE
                      fig.align = "center",
                      echo = FALSE      # incluir chunk en output
                      )
options(scipen=999) # notacion cientifica
rm(list=ls())       # limpiar workspace
options(knitr.kable.NA = '') # NA en kable = ''
```
\pagestyle{fancy}
\fancyhead[R]{ \thepage}
\fancyhead[L]{Research Article Draft}
# Introduction

[~500 words]

## Class divide in the economic domain

The link between social class and political attitudes constitutes one of the most relevant findings in sociology. Theoretically, studies in the social mobility field have defined classes as structural positions given their difference in labor market situations distinguished by employment relations, skills level, and authority within the workplace, which traditionally has been represented by occupations instead of relative income or educational credentials [@Erikson1992;@Erikson1979]. Besides the social stratification studies, political sociologists have systematically demonstrated that classes trace distinctions in other domains of social life as well, including ideological preferences [@evans_decline_1993], cultural consumption behavior [@Chan2007] and attitudes toward redistribution and welfare [@Svallfors2006]. Regarding the latter, during the last decades, the research on class and political attitudes has developed more extensively regarding the economic domain, addressing the economic interests of classes with an emphasis on redistribution, social security, and social service functions of the welfare state [@lindh_class_2020, p. 421]. 

Critical views have suggested that economic development and modernization have eroded the link between social class and political attitudes. Around the debate of the "death of class" thesis, it has been argued that due to the increase of material welfare in post-industrial societies, the distributive struggle between classes has been waning, resulting in the salience of other group identities as more relevant, reorienting the political debate towards post-material concerns [@clark_declining_1993;@inglehart_culture_1990]. However, although modernization has led to changes in citizens' concerns, it is feasible to argue that the distributive struggle between classes has not been wholly displaced but cohabits with people's attitudes in the sociocultural domain [@lindh_class_2020]. Empirically, the claimed blurring of the class cleavage has not been strongly supported, whereas class divides indeed shape attitudes in the economic domain in general, such as perceived class conflicts [@perez_class_2023;@edlund_democratic_2015], attitudes toward market-based inequality [@andersen_preferences_2018;@lindh_public_2015], and redistributive preferences in particular [@curtis_how_2015;@langsaether_more_2020;@brooks_why_2010]. In the broader context, scholars have identified two competing mechanisms to elucidate the relationship between social class and attitudes towards inequality, commonly referred to as the self-interest and normative value-driven motives perspectives.

Theoretically, self-interest has served as the canonical model to explain redistributive preferences, where the material interests of those with the most to gain from redistribution conflict with those in positions of greater material well-being [@meltzer_rational_1981]. Similarly, extensions of the model have paid more attention to occupation as a representation of potential unemployment risk, arguing that social classes with greater risk exposure are prone to support redistributive policies as they reduce labor market insecurity [@rehm_risks_2009]. Differently, part of the scholarly discussion has stressed the role of cognitive processes in preference formation for understanding why, even in contexts of greater economic disparities, the lower classes oppose redistributive policies, arguing that misinformation plays a crucial role in linking material conditions with the redistributive policies [@bartels_homer_2005;@druckman_preference_2000]. Similarly, it has been argued that contexts of greater segregation affect the inferences about social and economic inequality [@mijs_inequality_2018] that ultimately contribute to the decoupling of material interests from economic preferences [@hvidberg_social_2020;@garcia-castro_changing_2022].  

As material interests might prevail in scarcity, normative or value-driven motivations emerge as an alternative explanation in preference formation when pure self-interest weakens. Thus, values such as self-transcendence or humanitarianism share a common spirit regarding their desire for equality and concern for human dignity, providing the common ground for altruism that ultimately motivates political preferences [@Kulin2013;@feldman_humanitarian_2001]. Despite the protagonism of values, these are indeed conditioned to risk exposure and resource availability, becoming salient under conditions of greater security against market risks and material prosperity and weakening under conditions of greater vulnerability and material precariousness [@Maldonadoetal2019]. Besides, it has been argued that occupations play a role in political attitude formation as they represent a set of specific skills and work logic that convey particular normative views [@oesch_redrawing_2006]. Thus, workspaces are conceived as socializing agents, considering that economically active people spend a substantial part of their time at work, including the social relations involved in providing goods and services. Hence, empirical evidence has suggested that technical and managerial professionals hold more conservative political views than occupations in interpersonal services [@oesch_electoral_2018;@lindh_class_2020]. Despite this, it is challenging to disentangle if the class differences are indeed explained through work experiences because it is feasible to argue that political preferences are the result of early-life internalization of social norms, which might be crystalized in career self-selection paths, which also might be influenced by the organizational hiring requirements for specific jobs [@kitschelt_occupations_2014].

While prior research has predominantly concentrated on examining the impact of social class through an individualistic lens, it is noteworthy that more attention needs to be devoted to understanding the role of social environments in class relations. This omission is particularly surprising given that class positions are fundamentally rooted in production relations that make them inherently relational, not only in their economic underpinnings but also in the power dynamics entwined within class conflicts [@wright_comparative_1989]. Besides, the normative basis of class relations introduces the relevance of the dimensions of solidarity and reciprocity, which have been argued to provide the moral basis of the legitimacy and popular support for welfare schemes [@mau_moral_2003]. Hereby, we argue that the scope of class analysis in attitude formation urges to be extended further to the individual or household situation, stressing the role of social ties and segregation as part of the socialization processes involved in the internalization of shared norms, class identity formation, conflict, and cooperation between classes.


- What is the role of class-based network segregation on redistributive preferences?

- Does economic inequality play a role in the relationship between class-based network segregation and redistributive preferences?

# Theoretical views on class, social networks, and redistributive preferences

[~3000 words]

## Class relations and social networks

As social relations within the occupational structure serve as the foundation for class analysis, widening our examination to other dimensions of social life, such as family dynamics, the development of friendships, and diverse social activities, can significantly enrich the scope of class analysis, particularly concerning its relevance in political attitudes formation. In principle, social stratification scholars have traditionally paid attention to _differentiation_ as the observable patterns of association given by the intersection of status and group membership, which can be understood as relational networks between positions [@blau_inequality_1977]. Empirically, homophily is one of the most significant findings in network research, whereas the likelihood of tie formation between individuals of similar characteristics is strong [@mcpherson_birds_2001; @lazarsfeld_friendship_1954] around demographic characteristics that trace friendship and family ties such as age and gender, as well as the segregated social environments alongside ethnicity and socioeconomic status [@bargsted_pautas_2020; @plaza_juntos_2022]. In addition, psychosocial factors such as socialization preferences and group dynamics also play a role in segregation processes, considering that isolated social environments decrease the interchange of world views, reinforcing in-group selectivity and value similarity [@visser_attitudes_2004;@homans_human_1951]. Nonetheless, without denying the relevance of socialization preferences, segregation has been predominantly explained through the lens of social differentiation in social activities as foci of contact opportunities that ultimately consolidate social networks [@feld_focused_1981; @mcpherson_network_2019]. From this scope, we argue that class relations not only represent resources-based distinctions, but patterns of sociability and exchange observed in the differentiation of social ties.

How classes are formed also provides insights regarding their social relations and sociability practices. In sociology, classes are understood as the everyday basis for action, given their labor market situation, that ultimately shapes their access to opportunities and resources [@weber_economy_1978]. According to habitus theory, classes incur a series of distinctive practices pivoted in cultural capital as the basis for mutual recognition [@bourdieu_distinction_1984]. Thus, classes seek to improve their position through the intergenerational transmission of resources, whereas similarities in friendship and family choices play a crucial role in the reproduction privilege. Additionally, while symbolic resources reinforce status distinctions, access to exclusive social activities in the upper class increases segregation at the expense of excluding the lower classes [@bourdieu_reproduccion_1981]. Ultimately, classes consolidate their demographic and cultural identities by mobilizing resources and shared sociability practices, ensuring intergenerational reproduction of their structural positions [@goldthorpe_sobre_1992]. Hence, the socialization of common lifestyles and shared worldviews facilitates the consolidation of social classes, whereas social resources embedded in social networks follow a pattern of accumulation and exclusion that ultimately leads to social segregation in both the upper and the lower bounds of the social structure.
 
Previous research on social class and tie formation has suggested that class permeability, understood as the formation of cross-class social ties, is traced differently across property and authority boundaries. Empirically, evidence from Western industrialized societies has shown that the formation of friendship ties between owners and the working class is very less likely, in contrast to the tendency of higher tie formation between supervisors and workers, suggesting that the authority dimension is much less permeable than authority-based boundaries given that class interests increase the social distance between proprietors and workers, while the intermediate class position of supervisors and contact frequency with workers make friendship tie formation more likely [@wright_relative_1992]. Similarly, evidence from Chile has shown that the higher permeability of the intermediate classes contrasts with the more homogeneous networks of the working class, suggesting that their limited resources and lower capacity to be socially engaged ultimately result in a lack of social resources that isolate them from others, while the upper class is less permeable and homogenous as a result of its tendency to self-selection as a practice that ultimately seeks to reproduce their privileged status positions [@otero_open_2021]. 

In contrast, cross-class embeddedness can be described regarding their social activities as well as through their changes across generations and time. In this regard, @pichler_social_2009 suggest that higher civic engagement in formal organizations increases the chances of bridging with diverse people among the upper class, in contrast to the working class, which tends to be more homogeneous in its civic engagement behavior. Similarly, other studies in Europe and South America have exhibited that the upper and intermediate classes hold increasingly diverse and prestigious social environments than the working classes, where being socially mobile does help to improve both dimensions in tie formation, whereas it does not equalize the weight of class background compared to those intergenerationally stable in the upper class [@cepic_how_2020;@carrascosa_class_2023]. Additionally, the few longitudinal evidence suggests that networks change following a dynamic of cumulative advantages regarding its composition in terms of prestige and diversity, where upper classes improve in both dimensions while the lower class shows more stability across the life course [@volker_social_2020]. 

Theoretically, these studies account for class-based network segregation as the degree of connectedness of an individual to different occupations is meant to represent social resources embedded in social networks vertically in the social structure [@lin_social_2007]. Drawing on this approach, the attempts of stratification scholars have been focused on how social connections are distributed across the class structure as a matter of social integration [@blau_inequality_1977]. Besides, _homogeneity_  is described as the lack of cross-class network ties. It is indeed conceptually more proximate to the homophily principle because it is anchored in ego’s class position. At the same time, _diversity_ is defined as the rate of dissimilar ties or simply the total ties to certain groups that do not necessarily count with a reference position to describe the network composition. Hence, both approaches provide a set of alternatives for the study of social class from a network perspective that has been increasingly discussed in the stratification literature as well as in the scholarly discussion about social class and political attitudes. 

## Network segregation and attitudes toward redistribution 

Apart from the individual approach to attitude formation, we argue that people also form their opinions in the economic domain based on their social relations. Despite the limited research on the link between social ties and attitudes in the economic domain, we identify two broad theoretical approaches that have discussed the role of the social environment in attitude formation.

On the one hand, one approach accentuates the role of beliefs about economic inequality rooted in social comparison processes as an explanation through which people form their opinions about redistributive policies [@condon_economic_2020]. This hypothesis can be traced to the studies on class images and perceived class conflicts [@Evans1992;@KelleyEvans1995]. Here, the initial argument is that people form their beliefs through the individual, family, friends, and coworkers experiences instead of the whole society, which is described as an availability heuristic that systematically biases inferences about inequality based on the homophily of reference groups [@Evans1992, p. 467]. From this perspective, how people infer the social world is linked to the degree of segregation in their immediate social environment, which influences the intensity and character of the information, ultimately shaping inequality perceptions [@mijs_is_2021]. Thus, experience sharing in conversations with socioeconomically diverse networks has been proven to contribute to the accuracy of the images of income and wealth inequalities compared to people in more isolated discussion networks [@summers_deliberating_2022]. Despite this, we argue that this research field has been more focused on the cognitive dimension of preference formation through inequality perceptions, either relying on surveys [e.g. @cansunar_who_2021;@garcia-castro_perceived_2022] or experimental manipulations [e.g. @Cruces2011;@becker_significant_2021], rather than empirically addressing the role of class segregation in social networks and its claimed influence on attitudes in the economic domain.

Differently, it has been suggested that the lens of a network approach provides a better picture of the class relations that nurture social norms and group identity [@kalmijn_social_2007, p. 550]. This claim resembles the fact that classes are characterized as collectivities with differences in their degrees of cohesion and solidarity, encompassing unequal status-based social interactions that are linked to the individual or household material well-being, cultural perspectives, and political preferences that structure the broader social experiences [@morris_attenuation_1996, p. 48]. Furthermore, these social relations contribute to the degree to which people reflect on the collective, so that in societies where opportunities for contact between people of different social classes can be detrimental to social integration. It has been argued that the social distance among groups widened by extreme inequality can be described as an _empathy gulf_ that comprises barriers to imagining others’ lifestyles and goods, which can also be expressed in spatially segregated interactions and isolated lifeworlds that provide meaning to social inequality experiences, resulting in doubts about worthiness when the lower classes contrast themselves with extravagant lifestyles from the upper classes, undermining feelings of social inclusion and cohesion [@sachweh_moral_2012]. Consequently, segregation drives the lives of others to become more distant and might have consequences on empathy and solidarity toward others, potentially leading to perceiving fellow citizens as strangers [@otero_lives_2022, p. 758]. Thus, we suggest that attitudes towards economic inequality may be influenced by the class situation of both the individual and network ties, resembling the core claims of intergroup contact theory [@pettigrew_metaanalytic_2006] so that social influence processes can either align or polarize attitudes according to the characteristic of groups as well as their contact opportunities given the degree of segregation. 

In recent years, we have seen an increase in this _network turn_ of class analysis, contributing to unraveling the relationship between the composition of social ties and attitudes toward inequality. In this regard, it has been shown that the degree of contact between different classes has shown that individuals of the lower class interact with those of upper classes may consider economic inequality as justifiable, whereas interactions with those of the lower class may prompt individuals from different social classes to question the fairness of income distribution [@vargassalfate_contact_2023]. Further evidence suggests that class-based contact diversity shortens the social distance between social classes, providing broader images about the living conditions of others, which can trigger greater concerns about inequality [@otero_power_2023] and undermine support for the market distribution of social services [@beck_explorando_2019]. 

In addition, studies that have more directly addressed the relationship between networks and social class have been consistent with previous findings. In this regard, evidence from Sweden suggests that people tend to assimilate their opinions according to their surrounding friends and acquaintances, which is reinforced by class homophily, where higher contacts in the managerial class negatively impact redistributive demands, while ties to the working class increase them arguing that “individuals take an impression from others and modify their attitudes accordingly” [@lindh_missing_2021, p. 698]. Similarly, additional socialization sources are cross-class ties through the family members. Family of origin is crucial for political socialization, where class interests and norms are nurtured in childhood and early adulthood. A study in the United States has shown that those tied to the upper class through parental relations are less supportive of redistribution and progressive taxation than those with working-class family backgrounds [@lee_consider_2023]. Besides, @paskov_crossclass_2022 argues that the class position of family and partners shapes preferences en tanto que los households share risk according to the class position of its members, where class homogeneity in ties polarizes preferences between lower and upper classes. In contrast, this gradient is blurred as heterogeneity increases according to the class position of the partner and family of origin.

In sum, the attitudes of individuals toward redistribution are influenced by their class-embedded social relations. Thus, the degree of class segregation in these networks plays a crucial role in shaping people's beliefs about income distribution and redistributive policies. Furthermore, the degree of contact between different classes and class homophily reinforces the assimilation of opinions according to surrounding friends and acquaintances. Therefore, we expect that being segregated in similar class environments polarizes own class interests, where segregated lower (upper) classes hold more (less) redistributive preferences. 

## Economic Inequality, segregation, and redistributive preferences

Income inequality can deteriorate social bonding and increase segregation.
 
* status anxiety among the upper-classes [@wilkinson_spirit_2010] and lower interpersonal trust  [@kragten_income_2017] 

* limited opportunities for social mobility consequently reduce class permeability in social networks [@otero_differences_2023] 


However, it has been discussed that inequality might also increase egalitarian preferences among those in better material situations. 

* concerns about the negative consequences (e.g. crime) of inequality on overall social welfare, particularly in high-status groups (Altruism) [@rueda_who_2019]
 
* convergence in egalitarian preferences in the upper-classes can be normatively grounded according to views about equality of opportunities (Procedural Justice) [@sachweh_why_2019]

\pagebreak 

# Overview

## At the micro-level:

- (H1) Being segregated in similar class environments polarize own class interests:
    * Segregated lower (upper) classes hold more (less) redistributive preferences

## At the macro-level:

- (H2a) Scenario A (booster): Inequality increases the influence of class segregation on redistributive preferences (polarization) 

- (H2b) Scenario B (mitigation): Inequality diminishes the influence of class segregation on redistributive preferences (consensus) 



# Data, variables and method

[1500 words]

# Results



# Discussion and conclusion


`r if (is_latex_output()) '
# References {-}
'`
```{=tex}
\scriptsize
\singlespacing
```


